---
output:
  pdf_document: default
  html_document: default
---

GRP78 expresion in childhood B-LLA
 
by Guadalupe Rivera Torruco 


This analysis correspond to multiparamatric cytometry bone marrow samples of childhood LLA de novo diagnosed. The stained samples were runned in  Cytoflex LX cytometer and pre-analysed in FlowJo v10.2. The cells analysed correspond to compesated Live CD45 cells.

```{r Suppress warning messages}
options(warn=-1)
```

```{r install and load R base packages}

packages <- c("knitr",
              "cowplot",
              "ggplot2",
              "readxl",
              "lattice",
              "uwot",
              "rstudioapi",
              "BiocManager",
              "knitr") 

for (i in packages){
  if(!is.element(i, .packages(all.available = TRUE))){
    install.packages(i)
  }
  library(i, character.only = TRUE)
}
```
```{r Select figure dimensions}
on.bioc <- TRUE

# Use fig.width = 7 for html and fig.width = 6 for pdf
fig.width <- ifelse(on.bioc, 8, 6)
knitr::opts_chunk$set(cache = 2, cache.path = "cache/",
                      warning = FALSE, message = FALSE, error = FALSE,
                      fig.path = "figure/", fig.width = fig.width)
options(width=75)
```

```{r install and load Bioconductor packages}
packages_bioconductor <- c("CATALYST",
                           "diffcyt",
                           "flowCore",
                           "scater",
                           "flowViz",
                           "BiocStyle")

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install(packages_bioconductor, force = TRUE, ask = FALSE)


```


```{r Multiload Bioconductor packages, message=FALSE}
#Bind the lists of packages and multi-load them
#all_packages <- c(packages, packages_bioconductor)

multi_library <- function(packages){
  for(i in packages){
    library(i, character.only = TRUE)
  }
}

multi_library(packages_bioconductor)
```

```{r Set working directory}
path <- "C:/Users/lupit/OneDrive/Analisis14feb2021" #path to files
setwd(path)
```

```{r Read Metadata file}
md <- "LLA_MO_GRP78_metadata.xlsx"
md <- read_excel(md)
head(data.frame(md))
```

```{r Read de FCS files}
#Read a flowset - multiple FCS files (when in the working directory)
#fs <- read.flowSet(path = 'C:User/bla/bla/fullFCS', pattern = ".fcs") <- when outside directory

flowset_LLA <- read.flowSet(path ="./Exports2", 
                            pattern = '.fcs')

#Compensation matrix
Spill <- flowset_LLA@frames[["LLA_MO_MELANI_HR.fcs"]]@description$`$SPILLOVER`



flowset_LLA.comp <- compensate(flowset_LLA, Spill)
```
We must create a csv or xlsx with the channel colnames in the FCS files and the antigen of our markers. i.e. channel FL1-A correspond to FITC channel, the antibody to recognize GRP78 is marked with FITC.
```{r Load panel data}
panel <- "LLA_MO_GRP78_panel_v3.xlsx"
panel <- read_excel(panel)
head(data.frame(panel))
```
We need to check if every fcs_colname is present in our flowset, if the
panel.xslx is correct
```{r Check prerequisites to build single cell exp (sce)}
all(panel$fcs_colname %in% colnames(flowset_LLA))

```
```{r specify levels for conditions & sample IDs to assure desired ordering}

md$condition <- factor(md$condition, levels = c("Standard risk", "High risk"))
md$sample_id <- factor(md$sample_id, levels = 
                             md$sample_id[order(md$condition)])
```
```{r Construct sce}
sce <- prepData(flowset_LLA, panel, md,
                md_cols = list(file = "file_name", id = "sample_id",
                               factors = c("condition", "LLA_id")))
```
After creating sce experiment we can see some diagnostic plots. If we want
to save it as pdf we can use:
pdf(file = "desired_name.pdf")
code_to_plot
dev.off()

Expression plot help us to see homogeneity between groups and see differences between them. 
```{r Diagnostic plots}
plotExprs(sce, color_by = "condition", assay = "exprs")
```
```{r Check number of cell per sample}

n_cells(sce)
```

```{r Plot number of cells, barplot}

plotCounts(sce, group_by = "sample_id", color_by = "condition")

```
```{r Multi-dimensional Scaling (MDS) plot}}
pbMDS(sce, color_by = "condition", label_by = "sample_id") 
```
```{r Plot-dendogram (Heatmap of the median marker expression)}

plotExprHeatmap(sce, scale = "last") 
```

```{r Marker ranking based on the non-redundancy score}

#pdf(file = "plotNRS.pdf")

plotNRS(sce, features = "type", color_by = "condition") #ERROR

#dev.off()

```
Following the exploratory analysis we will start clustering with FlowSOM and ConsensusClustering tools. We might need to set a seed, we can try different numbers:

# example: set.seed(1234)
#          set.seed(3332)

when not provided with a seed, overwrite our seed using the current system time:
set.seed(as.numeric(Sys.time()))

```{r Cell population identification with FlowSOM and ConsensusClusterPlus}
set.seed(6666)
sce <- cluster(sce, features = "type",
               xdim = 10, ydim = 10, maxK = 10, seed = 6666)
```

```{r Plot the entire marker expression distribution in each cluster}

plotExprHeatmap(sce, features = "type", 
                by = "cluster_id", k = "meta10", 
                bars = TRUE, perc = TRUE) 
```

```{r Plotting abudance of each cluster by sample id}

plotAbundances(sce, k = "meta10", by = "sample_id")

```
```{r plot Abundances}

plotAbundances(sce, k = "meta10", by = "cluster_id")

```

t-SNE and UMAP based on the arcsinh-transformed expression of the 6 lineage 
markers in the cells from the dataset. From each of the 6 samples, 1000 cells were randomly selected. Cells are colored according to the 10 cell subpopulations obtained with FlowSOM after the metaclustering step with ConsensusClusterPlus.

```{r Visual representation with tSNE}
set.seed(6666)
sce <- runDR(sce, "TSNE", cells = 1000, features = "type")
```

```{r tSNE stratified by sample}
plotDR(sce, "TSNE", 
       color_by = "meta10", 
       facet_by = "sample_id")
```

```{r tSNE stratified by condition}
plotDR(sce, "TSNE", 
       color_by = "meta10", 
       facet_by = "condition")

```

```{r Color by single marker, split by sample}
plotDR(sce, "TSNE", color_by = "GRP78", 
       facet_by = "condition", 
       ncol = 4,
       a_pal = rev(hcl.colors(10, "Spectral")))
```

```{r Color by all markers using custom palette}

plotDR(sce, "TSNE", color_by = rownames(sce), ncol = 6, 
       a_pal = rev(hcl.colors(10, "Spectral")))
```

```{r Color by scaled expression for set of markers, split by condition}
plotDR(sce,
       "TSNE",
       scale = TRUE, 
       facet_by = "condition",
       color_by = sample(rownames(sce), 6),
       a_pal = rev(hcl.colors(10, "Spectral")))
```

```{r Visual representation with UMAP}
set.seed(6666)
sce <- runDR(sce, "UMAP", cells = 1000, features = "type")
```

```{r UMAP stratified by sample}
plotDR(sce, "UMAP", 
       color_by = "meta10", 
       facet_by = "sample_id")
```

```{r UMAP stratified by condition}
plotDR(sce, "UMAP", 
       color_by = "meta10", 
       facet_by = "condition")

```

```{r UMAP Color by single marker, split by sample}
plotDR(sce, "UMAP", color_by = "GRP78", 
       facet_by = "condition", 
       ncol = 4,
       a_pal = rev(hcl.colors(10, "Spectral")))
```

```{r UMAP Color by all markers using custom palette}

plotDR(sce, "UMAP", color_by = rownames(sce), ncol = 6, 
       a_pal = rev(hcl.colors(10, "Spectral")))
```

```{r UMAP Color by scaled expression for set of markers, split by condition}
plotDR(sce,
       "UMAP",
       scale = TRUE, 
       facet_by = "condition",
       color_by = sample(rownames(sce), 6),
       a_pal = rev(hcl.colors(10, "Spectral")))
```

---
title: LLA_MO_GRP78_.R
author: Guadalupe Rivera Torruco
date: '2021-07-16'
output:
  pdf_document:
    latex_engine: xelatex
sansfont: Times New Roman

---
